---
title: "Composers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Composers}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`heyahmama` does not have all the composers to all the songs yet.
This vignette shows the progress to this.

```{r}
library(heyahmama)
library(testthat)
```

First we get all composers:

```{r}
composers <- get_composers()
knitr::kable(head(composers))
```

## See all composers

Songs are written by 1 or more composers.

Here are all different composers:

```{r}
unique_composers <- sort(unique(composers$composer))
n_composers <- 11
expect_equal(n_composers, length(unique_composers))
```

There are `r n_composers` different composers.

## Count songs with missing composers

Now, we search through all songs
and if these are mentioned in the
table `composers`:

```{r}
t <- tibble::tibble(
  song_title = sort(get_songs()$song_title)
)
t$has_composer <- t$song_title %in% composers$song_title
knitr::kable(head(t))
```

And the missing songs only:

```{r}
knitr::kable(t[t$has_composer == FALSE,])
```

In a plot:

```{r}
tally <- dplyr::count(t, has_composer)
ggplot2::ggplot(
  tally,
  ggplot2::aes(x = has_composer, y = n)
) + ggplot2::geom_col()
```

## Count compositions

Songs have different compositions of composers.
For example, 'Heyah mama' has three composers:

```{r}
composers[composers$song_title == "Heyah mama", ]$composer
```

The amount of composers differs too.

Here we count how many composers a song has:

```{r}
n_composers_per_song <- composers |> 
  dplyr::group_by(song_title) |>
  dplyr::count(name = "n_composers")
knitr::kable(head(n_composers_per_song))
```

Here we determine the frequency of the amount of composers:

```{r}
n_composers_frequency <- n_composers_per_song |>
  dplyr::group_by(n_composers) |>
  dplyr::count(name = "frequency")
knitr::kable(n_composers_frequency)
```

Most songs are written by three composers.
Which three?

```{r}
songs_with_three_composers <- n_composers_per_song[n_composers_per_song$n_composers == 3, ]$song_title

composers_working_in_trio <- unique( composers[composers$song_title %in% songs_with_three_composers, ]$composer)

sort(composers_working_in_trio)
```

Ouch, there are multiple composers working in a trio.

Instead of focusing on trios only, lets invesitigate
the frequency of all groups.

```{r}
composer_names <- sort(unique(composers$composer))
song_titles <- sort(unique(composers$song_title))
n_cols <- length(composer_names) + 1
n_rows <- length(song_titles)
col_names <- c("song_title", composer_names)
m <- matrix(nrow = n_rows, ncol = n_cols)
colnames(m) <- col_names
t <- tibble::as_tibble(m)
t$song_title <- song_titles

for (col_index in seq(2, n_cols)) {
  message(names(t)[col_index])
  composer_name <- names(t)[col_index]
  songs_with_this_composer <- composers[composers$composer == composer_name, ]$song_title
  t[, col_index] <- t$song_title %in% songs_with_this_composer
}
knitr::kable(head(t))
```

Now assign a group ID for each formation:

```{r}
bools_to_int <- function(bools) {
  exponent <- 0
  sum <- 0
  for (b in rev(bools)) {
    sum <- sum + (as.integer(b) * (2 ^ exponent))
    exponent <- exponent + 1
  }
  sum
}
expect_equal(0, bools_to_int(bools = c()))
expect_equal(0, bools_to_int(bools = FALSE))
expect_equal(0, bools_to_int(bools = rep(FALSE, 10)))
expect_equal(1, bools_to_int(bools = TRUE))
expect_equal(2, bools_to_int(bools = c(TRUE, FALSE)))
expect_equal(3, bools_to_int(bools = rep(TRUE, 2)))
expect_equal(4, bools_to_int(bools = c(TRUE, FALSE, FALSE)))

t$group_id <- NA
for (i in seq_len(n_rows)) {
  bools <- as.logical(t[i, 2:n_cols])
  t$group_id[i] <- bools_to_int(bools)
}

knitr::kable(head(t |> dplyr::select(song_title, group_id)))
```

Let's count the group IDs:

```{r}
group_ids_count <- t |> dplyr::select(song_title, group_id) |> dplyr::count(group_id)
knitr::kable(group_ids_count)
n_composer_groups <- nrow(group_ids_count)
expect_equal(5, n_composer_groups) 
```

So, although there are many composers,
there are only `r n_composer_groups` composer
groups.

Which are the names of those in the most
frequent composer group?

First get the group ID of the most prolific group:

```{r}
most_prolific_group_id <- group_ids_count[group_ids_count$n == max(group_ids_count$n), ]$group_id
most_prolific_group_id
```

```{r}
first_row_of_most_prolific_group <- t[t$group_id == most_prolific_group_id, ][1, ]
names_of_most_prolific_group <- names(first_row_of_most_prolific_group)[as.logical(first_row_of_most_prolific_group[1, ] == TRUE)]
expected_names <- c("Alain Vande Putte", "Miguel Wiels",      "Peter Gillis")
expect_equal(expected_names, names_of_most_prolific_group)
names_of_most_prolific_group
```

For which K3 formations have they written?

```{r}
songs_by_most_prolific_group <- t[t$group_id == most_prolific_group_id, ]$song_title

songs_on_cds <- dplyr::select(get_songs(), song_title, cd_title)
songs_on_cds_by_most_prolific_group <- songs_on_cds[
  songs_on_cds$song_title %in% songs_by_most_prolific_group,
]

cd_formations <- dplyr::select(get_cds(), cd_title, formation)

songs_per_formation_from_most_prolific_group <- dplyr::select(
  merge(songs_on_cds_by_most_prolific_group, cd_formations),
  song_title,
  formation
)
song_count_per_formation_from_most_prolific_group <-
  dplyr::count(
  songs_per_formation_from_most_prolific_group |> 
    dplyr::group_by(formation)
)

knitr::kable(song_count_per_formation_from_most_prolific_group)
```

Note that this most prolific group only wrote for the
first three K3 formations.
