---
title: "demo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{demo}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`heyahmama` is an R package containing data regarding
K3, the popular Flemish band. It provides
functions to easily access and work with that data.
The package is named after the K3 song 'Heyah Mama',
which is the first album CD of the band.

To use `heyahmama`, first load the package:

```{r setup}
library(heyahmama)
```

Showing the first 6 CDs:

```{r}
knitr::kable(head(get_cds()))
```

Showing the first 2 songs with its writers:

```{r}
knitr::kable(head(get_writers()))
```

We can use that data to plot how much songs were
written by each writer:

```{r}
writers <- get_writers()
writer_tally <- dplyr::count(writers, writer_name)
ggplot2::ggplot(
  data = writer_tally, 
  ggplot2::aes(x = writer_name, y = n)
) + ggplot2::geom_col() + 
  ggplot2::theme(
    axis.text.x = ggplot2::element_text(
      angle = 90,
      vjust = 0.5,
      hjust = 1
    )
  )
```

Showing all the band formations:

```{r}
knitr::kable(get_formations())
```

Counting which singers where in how many formations:

```{r}
formations <- get_formations()
member_tally <- dplyr::count(formations, member)
ggplot2::ggplot(
  data = member_tally, 
  ggplot2::aes(x = member, y = n)
) + ggplot2::geom_col() + 
  ggplot2::theme(
    axis.text.x = ggplot2::element_text(
      angle = 90,
      vjust = 0.5,
      hjust = 1
    )
  )
```

Showing the first 6 songs:

```{r}
knitr::kable(head(get_released_songs()))
```

Combining these two tables, we can calculate how many
songs are released by each formation:

```{r}
songs <- get_released_songs()
cds <- get_cds()
songs_and_cds <- merge(songs, cds)
song_per_formation <- dplyr::count(songs_and_cds, formation)
ggplot2::ggplot(
  data = song_per_formation, 
  ggplot2::aes(x = formation, y = n)
) + ggplot2::geom_col() + 
  ggplot2::theme(
    axis.text.x = ggplot2::element_text(
      angle = 90,
      vjust = 0.5,
      hjust = 1
    )
  )
```

